An investigation of the clinical utility of preoperative cardiopulmonary exercise testing in patients undergoing major pancreatic surgery.

### 2016-02-13 Presentations/viva_presentation_short.tex
Viva Presentation
- Down to 35 slides
- Can remove a lot from the intro

\begin{frame}
	\frametitle{Parameters measured during CPET }
	\framesubtitle{ Section 1.6 Page 25}
	\begin{itemize}
		\item
		\item
		\item
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Patients and Methods}
	\begin{itemize}
		\item Patients undergoing pancreaticoduodenectomy at Glasgow Royal Infirmary
		\item August 2008 - August 2012
		\item Prospectively maintained MS Access database
		\item Preoperative data - Bloods, co-morbidity including POSSUM Physiology Score, stent information
		\item Cardiopulmonary exercise testing
		\item Prospective recording of complications using ISGPS or Clavie-Dindo classification
		\item Serial measurements of biochemical parameters during first postoperative week
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{CPET, Jaundice and Preoperative Pathophysiology}
	\framesubtitle{What were the preoperative patient factors that were associated with a low $\dot{V}_{O_2}$?}
	Factors associated with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min
	\begin{table}
		\begin{tabular}{l l l l}
			& HR   & 95\% CI    & p     \\ \hline
			Female sex & 3.75 & 1.57-8.95  & 0.003 \\
			BMI$>$25   & 3.65 & 1.61-8.26  & 0.002 \\
			Cancer     & 4.02 & 1.33-12.16 & 0.014 \\
			CRP$>$10   & 2.98 & 1.29-6.86  & 0.010
		\end{tabular}
	\end{table}

	Factors associated with $\dot{V}_{O_2}$Peak $<$ 16ml/kg/min
	\begin{table}
		\begin{tabular}{l l l l}
			& HR   & 95\% CI   & p        \\ \hline
			Female sex  & 7.57 & 3.09-18.5 & $<$0.001 \\
			BMI$>$25    & 2.57 & 1.18-5.63 & 0.018    \\
			Haemoglobin & 2.43 & 1.05-5.63 & 0.038
		\end{tabular}
	\end{table}
\end{frame}

@@

	\column[c]{0.5\textwidth}		
	{\scriptsize			
	Severity of systemic inflammation proportionate to severity of jaundice. \\
	\medskip

	Obstructive jaundice associated with electrolyte abnormalities.\\
	\medskip

	Jaundiced patients were more likely to be anaemic, with lower haematocrit and lower mean corpuscular volume.\\
	\medskip

	No significant difference in the preoperative renal function, prothrombin time and white cell count between jaundiced and non-jaundiced patients.\\
	}

@@

\begin{frame}
	\frametitle{Chapter 4}
	\framesubtitle{CPET and Body Composition - Page 97-122 }
	\begin{block}{Title}
		An investigation into the relationship between cardiopulmonary exercise testing and body composition in patients undergoing pancreaticoduodenectomy.
	\end{block}
\end{frame}

\begin{frame}
	\frametitle{What causes a low $\dot{V}_{O_2}$? }
	\framesubtitle{Section 4.1.2 Page 98-99 }
	\begin{itemize}
		\item Low $\dot{V}_{O_2}$ has been universally ascribed to low aerobic fitness due to inadequate cardio-pulmonary reserve
		\item Occassionally anaemia, peripheral vascular disease or even mitochondrial diseases may play a role
		\item $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak are most commonly used parameters
		\item Both are normalised by dividing using patient's total body weight
		\item But, during exercise $\dot{V}_{O_2}$ is a function of skeletal muscle mass
		\item This can result in 'spurious correlation' unfairly penalising obese subjects
		\item Most patients in this cohort did not have overt cardiac/respiratory disease that could explain the low $\dot{V}_{O_2}$.
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Body composition estimation}
	\framesubtitle{Figure 4.1 Page 103}
	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{bc_gimp}
		\label{fig:bc_gimp}
	\end{figure}
\end{frame}

@@

\begin{frame}
	\frametitle{Chapter 5}
	\begin{block}{Title}
		An investigation into the relationship between preoperative clinico-pathological characteristics and post-operative systemic inflammatory response in patients undergoing pancreaticoduodenectomy.
	\end{block}
\end{frame}

\begin{frame}
	\frametitle{Background}
	\framesubtitle{Section 5.1 Page 124}
	\begin{itemize}
		\item Perioperative systemic inflammation influences short and long-term outcomes
		\item Elevated preoperative systemic inflammation is associated with postoperative complications in colorectal surgery, after oesophagectomy, liver resections and cardiac surgery. 
		\item Early exaggerated postoperative systemic inflammatory response is also associated with increased infective complications
		\item In patients undergoing pancreatic surgery, preoperative inflammatory state is further influenced by:
		\begin{itemize}
			\item Obstructive jaundice
			\item Cholangitis
			\item ERCP and stenting
			\item Post-ERCP / EUS related complications
		\end{itemize}
		\item Obesity is associated with chronic inflammation
		\item Impact of aerobic capacity on postoperative inflammation not studied before
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Preop inflammation vs. Postop CRP}
	\framesubtitle{Figure 5.3 Page 142}
	\begin{figure}
		\includegraphics[width=\textwidth]{../Figures/sirs_crp_crp_alb}
	\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Obstructive jaundice vs. Postop CRP}
	\framesubtitle{ }
	\begin{figure}
		\includegraphics[width=0.7\textwidth]{../Figures/sirs_bil_crp}
	\end{figure}

@@

\section[Chapter 6]{Postoperative CRP and complications}
\begin{frame}
	\frametitle{Chapter 6}
	\framesubtitle{Postoperative CRP and complications}
	\begin{block}{Title}
		An investigation into the relationship between postoperative systemic inflammation and complications after pancreaticoduodenectomy.
	\end{block}
\end{frame}

@@

% Discuss the results in the context of a patient.
% 63 yr old female with BMI of 35 presents with painless OJ
% PMH - HTN, T2DM, smoked 10/day for 40 yrs
% Bloods: Bil 180, CRP 22, Alb 30
% Investigations reveal an operable tumour of the head of the pancreas
% How would this patient have been managed 10 years ago?
\section{Case study}
\begin{frame}
	\frametitle{A typical patient}
	\begin{itemize}
		\item 63 yr old female with BMI of 35 presents with painless OJ
		\item PMH - HTN, T2DM, smoked 10/day for 40 yrs
		\item Bloods: Bil 180, CRP 22, Alb 30
		\item Imaging: Resectable tumour of head of pancreas
		\item How would this patient have been managed 10 years ago?
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{10 years ago...}
	\begin{itemize}
		\item Non-dynamic measures of fitness - ECG, PFTs, ASA, ECHO
		\item Routine biliary drainage
		\item No objective methods to predict complications other than pancreatic leak
		\item Impact of postoperative SIRS on complications - no clinical utility
		\item Limited ability to identifying patients at risk of complications
		\item ERAS principles not universally applied
		\item No prehabilitation
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Now... }
	\begin{itemize}
		\item CPET in all or selected cases (evidence from other patient groups)
		\item Routine biliary drainage no longer indicated
		\item Enhanced recovery more widespread
		\medskip
		\hrule
		\medskip
		\item Reluctance to perform CPET in jaundiced patients
		\item No recognised thresholds for CPET parameters
		\item No clear thresholds for post-op inflammatory markers in predicting complications
		\item Poor understanding of impact of obesity on preop function and postop outcomes
		\item No prehabilitation
	\end{itemize}
\end{frame}

%------------------------------------------------

\begin{frame}[fragile] % Need to use the fragile option when verbatim is used in the slide
\frametitle{Verbatim}
\begin{example}[Theorem Slide Code]
\begin{verbatim}
\begin{frame}
\frametitle{Theorem}
\begin{theorem}[Mass--energy equivalence]
$E = mc^2$
\end{theorem}
\end{frame}\end{verbatim}
\end{example}
\end{frame}

%------------------------------------------------

>>>>

@@

@@

@@

		\column[c]{0.5\textwidth}		
		{\scriptsize			
			Severity of systemic inflammation proportionate to severity of jaundice. \\
			\medskip
			Obstructive jaundice associated with electrolyte abnormalities.\\
			\medskip
			Jaundiced patients were more likely to be anaemic, with lower haematocrit and lower mean corpuscular volume.\\
			\medskip
			No significant difference in the preoperative renal function, prothrombin time and white cell count between jaundiced and non-jaundiced patients.\\
		}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{What preop factors were associated with low $\dot{V}_{O_2}$?}
	\framesubtitle{Multivariate binary logistic regression analysis}
	Factors associated with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min
	\begin{table}
		\begin{tabular}{l| l l l}
			& HR   & 95\% CI    & p     \\ \hline
			Female sex & 3.75 & 1.57-8.95  & 0.003 \\
			BMI$>$25   & 3.65 & 1.61-8.26  & 0.002 \\
			Cancer     & 4.02 & 1.33-12.16 & 0.014 \\
			CRP$>$10   & 2.98 & 1.29-6.86  & 0.010
		\end{tabular}
	\end{table}
	\hrule
	\medskip
	Factors associated with $\dot{V}_{O_2}$Peak $<$ 16ml/kg/min
	\begin{table}
		\begin{tabular}{l| l l l}
			& HR   & 95\% CI   & p        \\ \hline
			Female sex  & 7.57 & 3.09-18.5 & $<$0.001 \\
			BMI$>$25    & 2.57 & 1.18-5.63 & 0.018    \\
			Haemoglobin & 2.43 & 1.05-5.63 & 0.038
		\end{tabular}
	\end{table}
\end{frame}

@@

\begin{frame}
	\frametitle{ CPET and Body Composition}
	\framesubtitle{What causes a low $\dot{V}_{O_2}$? }
	\begin{itemize}
		\item Low $\dot{V}_{O_2}$ has been universally ascribed to low aerobic fitness due to inadequate cardio-pulmonary reserve
		\item Occassionally anaemia, peripheral vascular disease or even mitochondrial diseases may play a role
		\item $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak are most commonly used parameters
		\item During exercise $\dot{V}_{O_2}$ is a function of skeletal muscle mass
		\item But, both are normalised by against patient's total body weight
		\item 'Spurious correlation' unfairly penalises obese subjects
		\item Most patients in this cohort did not have overt cardiac/respiratory disease that could explain the low $\dot{V}_{O_2}$.
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Body composition differences between the sexes}
	\framesubtitle{Significant differences in skeletal muscle area across all BMIs.}
	\begin{figure}
		\centering
		\includegraphics[height=0.95\textheight]{../Figures/bc_gender_bmi_pie}
		\label{fig:bc_gimp}
	\end{figure}
\end{frame}

@@

\begin{frame}
	\frametitle{Impact of perioperative systemic inflammation?}
	\framesubtitle{Inflammation before and after surgery affects short and long term outcomes}
	\begin{itemize}
		\item Perioperative systemic inflammation influences short and long-term outcomes
		\item Elevated preoperative systemic inflammation is associated with postoperative complications in colorectal surgery, after oesophagectomy, liver resections and cardiac surgery. 
		\item Early exaggerated postoperative systemic inflammatory response is also associated with increased infective complications
		\item In patients undergoing pancreatic surgery, preoperative inflammatory state is further influenced by:
		\begin{itemize}
			\item Obstructive jaundice
			\item Cholangitis
			\item ERCP and stenting
			\item Post-ERCP / EUS related complications
		\end{itemize}
		\item Obesity is associated with chronic inflammation
		\item Impact of aerobic capacity on postoperative inflammation not studied before
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Preop inflammation vs. Postop CRP}
	\framesubtitle{Post op CRP trends are influenced by both preop CRP and preop albumin.}
	\begin{figure}
		\includegraphics[width=\textwidth]{../Figures/sirs_crp_crp_alb}
	\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Obstructive jaundice vs. Postop CRP}
	\framesubtitle{Obstructive jaundice is associated with an attenuated inflammatory response.}
	\begin{figure}
		\includegraphics[width=0.7\textwidth]{../Figures/sirs_bil_crp}
	\end{figure}

@@

\section[Chapter 6]{Postoperative CRP and complications}

@@

% Discuss the results in the context of a patient.
% 63 yr old female with BMI of 35 presents with painless OJ
% PMH - HTN, T2DM, smoked 10/day for 40 yrs
% Bloods: Bil 180, CRP 22, Alb 30
% Investigations reveal an operable tumour of the head of the pancreas
% How would this patient have been managed 10 years ago?
\section{Case study}
\begin{frame}
	\frametitle{A case study}
	\framesubtitle{How does this work influence decision-making?}
	\begin{itemize}
		\item 63 yr old female with BMI of 35 presents with painless OJ
		\item PMH - HTN, T2DM, smoked 10/day for 40 yrs
		\item Bloods: Bil 180, CRP 22, Alb 30
		\item Imaging: Resectable tumour of head of pancreas
		\medskip
		\item Should this patient undergo CPET?
		\item Should this patient undergo biliary drainage?
		\item Should we wait until resolution of jaundice before CPET?
		\item How can the CPET results be used to influence her care?
		\item How can the postoperative inflammatory markers be used to manage her care?
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{A case study}
	\begin{description}
	{\scriptsize 	
		\item [Ch 2] CPET may allow identifying if she is at increased risk of complications, prolonged hospital stay and non-receipt of adjuvant therapy. 
					\\ CPET should not be used on its own to decline surgery. 
					\hrule
		\item [Ch 3] CPET can be performed in the presence of jaundice. 
					\\ Jaundice does not adversely affect dynamic cardiopulmonary response to and this cannot be a justification for biliary drainage. 
					\\ Taken together with evidence from trials on preperative biliary drainage, she should proceed direct to surgery.
						\hrule
		\item [Ch 4] Interpret low $\dot{V}_{O_2}$ in her case with caution. 
					\\ Her aerobic capacity may be better than that suggested by her test results. 
					\\ Beware of \textit{spurious correlation} with total body weight.
						\hrule
		\item [Ch 5] Jaundice and hypo-albuminemia attenuate post-op inflammatory response.
					\\ However, this does not translate into increased complications. 
					\\ Aerobic capacity does not affect post-op inflammatory response.
						\hrule
		\item [Ch 6] If her drain amylase on D3 is normal and D3 CRP$<$178 or D4 CRP$<$125, she is likely to have an uneventful postoperative course and must continue on 'enhanced-recovery' pathway.
				\\ If she develops a POPF, the early inflammatory response does not predict outcome.
	}
	\end{description}
\end{frame}

\begin{frame}
	\frametitle{Future directions}
	\begin{itemize}
		\item Assessing impact of prehabilitation
		\item Assessing impact of neo-adjuvant treatments
		\item Does CPET predict long-term survival?
		\item Improvements in aerobic capacity and body composition through nutritional supplementation and physical exercise
		\item Correction of anaemia through preoperative iron supplementation
		\item Early recognition and aggressive management of complications
		\item Assessing suitability for 'fast-track' to surgery and 'enhanced-recovery'
	\end{itemize}
